--- 
title: "New York Housing"
author: "Winnie Gao, Lingxuan Gu, Wenjie Xie"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
---
--- 
title: "New York Housing"
author: "Winnie Gao, Lingxuan Gu, Wenjie Xie"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
---
```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```

# Introduction


<!--chapter:end:index.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Data sources


The data sets we are going to use in our analyses come from NEW YORK Open Data, a popular website with various types of official data sets published by New York City agencies. To explore the rental apartments in NYC, we need to know some basic information about buildings, such as the time the buildings start to be built and completed, statistics about different apartment types, etc. Our first data set is Housing New York Units by Building, in which multiple information related to New York apartment information from 2014/01/01 to 2020/09/30 is collected: building id, construction type and counting of apartments with different number of bedrooms and number of units in the rent that are affordable to different earning populations. Since we also consider to explore spatial distribution about NYC apartments, location information, such as census tract, zip code, longitude and latitude, are also included in the data set. The type of variables are mostly number, basic text and date/time. There are 4,656 rows of building records in total and each row is a Housing Preservation and Development(HPD) building. The data type of features can be numeric, basic text and date.

Meanwhile, the pricing of the apartments in NYC is also a hot topic we want to explore. Unfortunately, apartment prices are not recorded in this data set. In this case, we are going to use another data set Local Law 44 Unit Income Rent which includes the number of bedrooms disaggregated by maximum allowable household income and initial legal and actual rents for each building in a Local Law 44 Housing Development Project. This data set do not include price information for all apartments, but, as a subset of all apartments, we can focus on this data set and make conclusions that can be generalized to all apartments in NYC. In this data set, building id, low/median/high initial and actual rent and total units number are recorded. Types of variables are number and plain text. There are in total 12597 rows of building records in this data set, each row is a bedroom Area Median Income(AMI) distribution.

The main issue of the second data set, Local Law 44 Unit Income Rent, is that there are a lot of missing values on the columns of initial and actual rent price. After observing the data set, we realize that all row with missing initial rents do not actual rents at the same time, and some rows only have actual rents missing Therefore, we can assume that the rows with missing values are those not rented, which can be disregarded during our research work. 


<!--chapter:end:02-data.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Data transformation

We download the two data sets mentioned in previous chapter directly from New York Open Data website in csv format so that the files can be easily read by R. After observing the data set, we notice that, in the data sets, NA are represented in different ways, using $"----"$ or $"\ "$(white space), which can not be directly identified as NA by r. So our first step is to replace these $“----”$ as NA. Recoding work is also done on some columns. In column MaxAllowableIncome” are same ranges for income such as "0%-30%" or "31%-40%" initially with levels in random orders. For future potential visualization work, it is better to reorder the levels into conceptually ascending order. Also, the numbers in the date set larger than 1000 are separated with commas. We need to remove these commas from the numbers and change their column types to numeric. Meanwhile, the date are in the format of MM-DD-YY in type of string. Data analysis and visualization work will be more convenient if we first change their type to date and use the format of YY-MM-DD.

After observing the patterns of NA values, we can confirm that the columns with around or more than 50% NAs are those columns that we do not really need for future analysis work. After removing these features, we can confirm that NAs in the rest columns can be roughly treated as randomly generated. For the rest of the data sets, all data points with NAs are removed. In other words, we also remove the entire row with NA appearing.a

Our data are recorded in two separate files with different features: one for rent related information and the other for geography and housing related information. Combining two data sets will allow us to have more insights on our information. Therefore, we also need to merge them by their building ID.

```{r}
proj = read.csv("Housing_New_York_Units_by_Building.csv", stringsAsFactors = TRUE, na.strings=c("----", "","NA"))
rent = read.csv("Local_Law_44_-_Unit_Income_Rent.csv", stringsAsFactors = TRUE, na.strings=c(" ", "","NA"))
data = merge(project, rent, by.x = "Building.ID", by.y = "BuildingID")

boxplot(data$MedianActualRent)
#head(sort(data$MedianActualRent, decreasing = T), 10)
```

From the boxplot of the `MedianActualRent` column, we could clearly see that there is one point with value larger than 10,000 that is far from all other data points. Thus, we could take this point as an outlier and remove it from the dataset.

<!--chapter:end:03-cleaning.Rmd-->


# Missing values

Placeholder



<!--chapter:end:04-missing.Rmd-->


# Results

Placeholder


## Analysis on Number of Projects by Time

<!--chapter:end:05-results.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Interactive component


```{r, echo=FALSE}
htmltools::includeHTML("inter.html")
htmltools::includeScript("interactive.js")
```

<!--chapter:end:06-interactive.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Conclusion


<!--chapter:end:07-conclusion.Rmd-->

